### 2022-03-22 manuscript/submission_review_2.tex
revisao 2, arquivo diff_2.pdf incluso

We present a fast equivalent layer method for processing large-scale magnetic data. 
We demonstrate that the sensitivity matrix associated with an equivalent layer
of dipoles can be arranged to a Block-Toeplitz Toeplitz-Block (BTTB) structure for the 
case in which observations and dipoles are aligned on a 
horizontal and regularly-spaced grid.
The product of a BTTB matrix and an arbitrary vector represents a discrete 
convolution and can be efficiently computed via 2D Fast Fourier Transform.
In this case, the matrix-vector product uses only the elements forming the first column
of the BTTB matrix, saving computational time and memory. 
Our convolutional equivalent layer method uses this approach to compute 
the matrix-vector products in the iterative conjugate gradient algorithm with the purpose 
of estimating the physical-property distribution over the equivalent layer for 
large data sets.
Synthetic tests with a mid-size $100 \times 50$ grid of total-field anomaly data
show a decrease of $\approx 10^4$ in floating-point operations and $\approx 25\times$ 
in computation runtime of our method compared to the classical approach of solving
the least-squares normal equations via Cholesky decomposition. 
Faster results are obtained for millions of data, showing drastic decreases in RAM usage
and runtime, allowing to perform magnetic data processing of large data sets on regular 
desktop computers. 
Our results also show that, compared to the classical Fourier approach, the magnetic
data processing with our method requires similar computation time, but produces significantly 
smaller border effects without using any padding scheme and also is more robust to 
deal with data on irregularly spaced points or on undulating observation surfaces.
A test with $1,310,000$ irregularly spaced field data over the Caraj{\'a}s Province, Brazil, 
confirms the efficiency of our method by taking $\approx 385.56$ seconds to estimate the physical-property
distribution over the equivalent layer and $\approx 2.64$ seconds to compute the upward 
continuation.

@@

Now we can define the structure of $\mathbf{A}$ in terms of its components 
$\mathbf{A_{\boldsymbol{\alpha\beta}}}$ (equation \ref{eq:A_expand}). To do this, 
we consider the particular case in which the observed total-field anomaly is located 
on an $N_{x} \times N_{y}$ 
regular grid of points spaced by $\Delta_{x}$ and $\Delta_{y}$ along the $x$- and $y$-directions,
respectively, on a constant vertical coordinate $z_{0}$. We also consider that the equivalent layer
is aranged such that one dipole is located right below each observation point, at a constant coordinate $z_{c}$.
In this case, the number of equivalent sources $M$ is equal to the number of data $N$ and, 
consequently, matrices $\mathbf{A}$ and $\mathbf{A_{\boldsymbol{\alpha\beta}}}$ become 
square ($N \times N$). 
Besides, the horizontal coordinates $x_{i}$ and $y_{i}$ of the observation points 
can be defined by
\begin{equation}
	x_{i} = x_{1} + \left[ k(i) - 1 \right] \, \Delta_{x}
	\label{eq:xi}
\end{equation}
and
\begin{equation}
	y_{i} = y_{1} + \left[ l(i) - 1 \right] \, \Delta_{y} \: ,
	\label{eq:yi}
\end{equation}
where $x_{1}$ and $y_{1}$ are the lower left corners coordinates for $x_{i}$ and $y_{i}$, respectively,
and $k(i)$ and $l(i)$ are integer functions defined according to the orientation
of the data grid (Figure \ref{fig:regular-grids}). 
For $x$-\textit{oriented grids}, the integer functions are given by
\begin{equation}
	k(i)  = i - \Bigg\lceil \frac{i}{N_{x}} \Bigg\rceil N_{x} + N_{x}
	\label{eq:k-x-oriented}
\end{equation}
and
\begin{equation}
	l(i) = \Bigg\lceil \frac{i}{N_{x}} \Bigg\rceil \: .
	\label{eq:l-x-oriented}
\end{equation}
For $y$-\textit{oriented grids}, the integer functions are given by
\begin{equation}
	k(i) = \Bigg\lceil \frac{i}{N_{y}} \Bigg\rceil
	\label{eq:k-y-oriented}
\end{equation}
and
\begin{equation}
	l(i) = i - \Bigg\lceil \frac{i}{N_{y}} \Bigg\rceil N_{y} + N_{y} \: .
	\label{eq:l-y-oriented}
\end{equation}
In equations \ref{eq:k-x-oriented}--\ref{eq:l-y-oriented}, $\lceil \cdot \rceil$ denotes the ceiling 
function \citep[e.g.,][ p. 67-68]{graham-etal1994}.
Equations \ref{eq:xi}--\ref{eq:l-y-oriented} can also be used to define the coordinates 
$x_{j}$ and $y_{j}$ of the equivalent sources, but with index $j$ instead of $i$.

@@

Table \ref{tab:RAM-usage} shows a comparison between the computer memory storage 
associated with each method. In the classical and standard CGLS methods, the whole 
matrix $\mathbf{A}$ (equation \ref{eq:A_expand}) has to be stored. For example, a dataset with 
$N = 10,000$ observation points has an associated sensitivity matrix $\mathbf{A}$ formed by 
$N^2 = 100,000,000$ elements and takes approximately $763$ Megabytes of memory (8 bytes per element). 
Using the same number of observation points $N = 10,000$, our method requires only 
$1.831$ Megabytes to store the first columns of the BCCB matrices
$\mathbf{C}_{\boldsymbol{\alpha\beta}}$ (equation \ref{eq:w_alpha_beta}) and 
$0.6104$ Megabytes to store the complex matrix $\mathbf{L}$ (equation \ref{eq:L}) 
(16 bytes per element). For a bigger dataset with $N = 1,000,000$, the amount of necessary RAM 
goes to $7,629,395$, $183.096$ and $61.035$ Megabytes, respectively.

@@

With $1,310,000$ observation points, it would be necessary $12.49$ Terabytes of RAM to store the full
sensitivity matrix with the classical method. 
In this case, our method uses only $59.97$ Megabytes, allowing regular desktop computers to be able 
to process this amount of data.

@@

Comparisons between the estimated physical-property distribution obtained with our method and 
the classical approach that solves the least-squares normal equations via Cholesky decomposition 
show similar results. 
The differences in total number of floating-point operations (flops), memory usage and computation 
time, however, are noticeable. 
For a mid-size grid of $100 \times 50$ points, the total number of flops is about four orders of
magnitude smaller than that required by the classical method. Besides, our method uses
less than $1\%$ of the RAM and takes about $3\%$ of the computation time associated with
the classical method in this case.% (TEM QUE CHECAR ESTAS VALORES)
Significantly better results can be obtained with larger data sets.

@@

While the classical equivalent-layer method would require $12.49$ Terabytes of RAM to store the full
sensitivity matrix associated with the irregular grid of $1,310,000$ observation points over the
Carajás Province, northern Brazil, our method requires only $59.97$ Megabytes.
When performed on a standard laptop computer with an Intel Core i7 7700HQ@2.8GHz processor in
single-processing and single-threading modes, the total times spent by our method to estimate the
physical-property distribution over the equivalent layer and to compute the upward continuation
of the $1,310,000$ magnetic observations over the Carajás province was approximately $385.56$ 
seconds and $2.64$ seconds.

@@

\tabl{RAM-usage}{This table shows the RAM usage (in Megabytes) for storing the whole 
	$N \times N$ matrix $\mathbf{A}$ (equation \ref{eq:A_expand}), the first columns of the BCCB
	matrices $\mathbf{C}_{\boldsymbol{\alpha\beta}}$ (equation \ref{eq:w_alpha_beta}) (both need 8 
	bytes per element) and the matrix $\mathbf{L}$ (equation \ref{eq:L}) (16 bytes per element).
	\label{tab:RAM-usage}}
{
	\begin{center}
		\begin{tabular}[]{|l|c|c|c|}
			\hline
			\textbf{$N$} & \textbf{$\mathbf{A}$} & \textbf{First columns of matrices $\mathbf{C}_{\boldsymbol{\alpha\beta}}$} & \textbf{$\mathbf{L}$}\\
			\hline 
			$100$ & 0.0763 & 0.0183 & 0.00610\\
			\hline
			$400$ & 1.22 & 0.0744 & 0.0248\\
			\hline
			$2,500$ & 48 & 0.458 & 0.1528\\
			\hline
			$10,000$ & 763 & 1.831 & 0.6104\\
			\hline
			$40,000$ & 12,207 & 7.32 & 2.4416 \\
			\hline
			$250,000$ & 476,837 & 45.768 & 15.3 \\
			\hline
			$500,000$ & 1,907,349 & 91.56 & 30.518 \\
			\hline
			$1,000,000$ & 7,629,395 & 183.096 & 61.035 \\
			\hline
		\end{tabular}
	\end{center} 
}
\clearpage

>>>>

We present a fast equivalent layer method for processing large-scale magnetic data. 
We demonstrate that the sensitivity matrix associated with an equivalent layer
of dipoles can be arranged to a Block-Toeplitz Toeplitz-Block (BTTB) structure for the 
case in which observations and dipoles are aligned on a 
horizontal and regularly-spaced grid.
The product of a BTTB matrix and an arbitrary vector represents a discrete 
convolution and can be efficiently computed via 2D Fast Fourier Transform.
In this case, the matrix-vector product uses only the elements forming the first column
of the BTTB matrix, saving computational time and memory. 
Our convolutional equivalent layer method uses this approach to compute 
the matrix-vector products in the iterative conjugate gradient algorithm with the purpose 
of estimating the physical-property distribution over the equivalent layer for 
large data sets.
Synthetic tests with a mid-size $100 \times 50$ grid of total-field anomaly data
show a decrease of $\approx 10^4$ in floating-point operations and $\approx 25\times$ 
in computation runtime of our method compared to the classical approach of solving
the least-squares normal equations via Cholesky decomposition. 
Faster results are obtained for millions of data, showing drastic decreases in computer memory usage
and runtime, allowing to perform magnetic data processing of large data sets on regular 
desktop computers. 
Our results also show that, compared to the classical Fourier approach, the magnetic
data processing with our method requires similar computation time, but produces significantly 
smaller border effects without using any padding scheme and also is more robust to 
deal with data on irregularly spaced points or on undulating observation surfaces.
A test with $1,310,000$ irregularly spaced field data over the Caraj{\'a}s Province, Brazil, 
confirms the efficiency of our method by taking $\approx 385.56$ seconds to estimate the physical-property
distribution over the equivalent layer and $\approx 2.64$ seconds to compute the upward 
continuation.

@@

Now we can define the structure of $\mathbf{A}$ in terms of its components 
$\mathbf{A_{\boldsymbol{\alpha\beta}}}$ (equation \ref{eq:A_expand}). To do this, 
we consider the particular case in which the observed total-field anomaly is located 
on an $N_{x} \times N_{y}$ 
regular grid of points spaced by $\Delta_{x}$ and $\Delta_{y}$ along the $x$- and $y$-directions,
respectively, on a constant vertical coordinate $z_{0}$. We also consider that the equivalent layer
is aranged such that one dipole is located right below each observation point, at a constant coordinate $z_{c}$.
In this case, the number of equivalent sources $M$ is equal to the number of data $N$ and, 
consequently, matrices $\mathbf{A}$ and $\mathbf{A_{\boldsymbol{\alpha\beta}}}$ become 
square ($N \times N$). 
Besides, the horizontal coordinates $x_{i}$ and $y_{i}$ of the observation points 
can be defined by
\begin{equation}
	x_{i} = x_{1} + \left[ k(i) - 1 \right] \, \Delta_{x}
	\label{eq:xi}
\end{equation}
and
\begin{equation}
	y_{i} = y_{1} + \left[ l(i) - 1 \right] \, \Delta_{y} \: ,
	\label{eq:yi}
\end{equation}
where $x_{1}$ and $y_{1}$ are the starting grid coordinates for $x_{i}$ and $y_{i}$, respectively,
and $k(i)$ and $l(i)$ are integer functions defined according to the orientation
of the data grid (Figure \ref{fig:regular-grids}). 
For $x$-\textit{oriented grids}, the integer functions are given by
\begin{equation}
	k(i)  = i - \Bigg\lceil \frac{i}{N_{x}} \Bigg\rceil N_{x} + N_{x}
	\label{eq:k-x-oriented}
\end{equation}
and
\begin{equation}
	l(i) = \Bigg\lceil \frac{i}{N_{x}} \Bigg\rceil \: .
	\label{eq:l-x-oriented}
\end{equation}
For $y$-\textit{oriented grids}, the integer functions are given by
\begin{equation}
	k(i) = \Bigg\lceil \frac{i}{N_{y}} \Bigg\rceil
	\label{eq:k-y-oriented}
\end{equation}
and
\begin{equation}
	l(i) = i - \Bigg\lceil \frac{i}{N_{y}} \Bigg\rceil N_{y} + N_{y} \: .
	\label{eq:l-y-oriented}
\end{equation}
In equations \ref{eq:k-x-oriented}--\ref{eq:l-y-oriented}, $\lceil \cdot \rceil$ denotes the ceiling 
function \citep[e.g.,][ p. 67-68]{graham-etal1994}.
Equations \ref{eq:xi}--\ref{eq:l-y-oriented} can also be used to define the coordinates 
$x_{j}$ and $y_{j}$ of the equivalent sources, but with index $j$ instead of $i$.

@@

Table \ref{tab:RAM-usage} shows a comparison between the computer memory storage 
associated with each method. In the classical and standard CGLS methods, the whole 
matrix $\mathbf{A}$ (equation \ref{eq:A_expand}) has to be stored. For example, a dataset with 
$N = 10,000$ observation points has an associated sensitivity matrix $\mathbf{A}$ formed by 
$N^2 = 100,000,000$ elements and takes approximately $763$ Megabytes of memory (8 bytes per element). 
Using the same number of observation points $N = 10,000$, our method requires only 
$1.831$ Megabytes to store the first columns of the BCCB matrices
$\mathbf{C}_{\boldsymbol{\alpha\beta}}$ (equation \ref{eq:w_alpha_beta}) and 
$0.6104$ Megabytes to store the complex matrix $\mathbf{L}$ (equation \ref{eq:L}) 
(16 bytes per element). For a bigger dataset with $N = 1,000,000$, the amount of necessary computer memory 
goes to $7,629,395$, $183.096$ and $61.035$ Megabytes, respectively.

@@

With $1,310,000$ observation points, it would be necessary $12.49$ Terabytes of computer memory to store the full
sensitivity matrix with the classical method. 
In this case, our method uses only $59.97$ Megabytes, allowing regular desktop computers to be able 
to process this amount of data.

@@

Comparisons between the estimated physical-property distribution obtained with our method and 
the classical approach that solves the least-squares normal equations via Cholesky decomposition 
show similar results. 
The differences in total number of floating-point operations (flops), memory usage and computation 
time, however, are noticeable. 
For a mid-size grid of $100 \times 50$ points, the total number of flops is about four orders of
magnitude smaller than that required by the classical method. Besides, our method uses
less than $1\%$ of the computer memory and takes about $3\%$ of the computation time associated with
the classical method in this case.% (TEM QUE CHECAR ESTAS VALORES)
Significantly better results can be obtained with larger data sets.

@@

While the classical equivalent-layer method would require $12.49$ Terabytes of computer memory to store the full
sensitivity matrix associated with the irregular grid of $1,310,000$ observation points over the
Carajás Province, northern Brazil, our method requires only $59.97$ Megabytes.
When performed on a standard laptop computer with an Intel Core i7 7700HQ@2.8GHz processor in
single-processing and single-threading modes, the total times spent by our method to estimate the
physical-property distribution over the equivalent layer and to compute the upward continuation
of the $1,310,000$ magnetic observations over the Carajás province was approximately $385.56$ 
seconds and $2.64$ seconds.

@@

\tabl{RAM-usage}{This table shows the computer memory usage (in Megabytes) for storing the whole 
	$N \times N$ matrix $\mathbf{A}$ (equation \ref{eq:A_expand}), the first columns of the BCCB
	matrices $\mathbf{C}_{\boldsymbol{\alpha\beta}}$ (equation \ref{eq:w_alpha_beta}) (both need 8 
	bytes per element) and the matrix $\mathbf{L}$ (equation \ref{eq:L}) (16 bytes per element).
	\label{tab:RAM-usage}}
{
	\begin{center}
		\begin{tabular}[]{|l|c|c|c|}
			\hline
			\textbf{$N$} & \textbf{$\mathbf{A}$} & \textbf{First columns of matrices $\mathbf{C}_{\boldsymbol{\alpha\beta}}$} & \textbf{$\mathbf{L}$}\\
			\hline 
			$100$ & 0.0763 & 0.0183 & 0.00610\\
			\hline
			$400$ & 1.22 & 0.0744 & 0.0248\\
			\hline
			$2,500$ & 48 & 0.458 & 0.1528\\
			\hline
			$10,000$ & 763 & 1.831 & 0.6104\\
			\hline
			$40,000$ & 12,207 & 7.32 & 2.4416 \\
			\hline
			$250,000$ & 476,837 & 45.768 & 15.3 \\
			\hline
			$500,000$ & 1,907,349 & 91.56 & 30.518 \\
			\hline
			$1,000,000$ & 7,629,395 & 183.096 & 61.035 \\
			\hline
		\end{tabular}
	\end{center} 
}
\clearpage
